1장 6절: 조현병의 자연경과

### 2022-03-18 article-01-04.Rmd
11장도 포함해서
2022-03-19

::: {#tab:dsm-5-specifier .table}
+-------------------------------------------------------------------------------------------+
| First episode, currently in acute episode (symptom criteria are fulfilled)                |
+-------------------------------------------------------------------------------------------+
| First episode, currently in partial remission (symptom criteria are not fully fulfilled)  |
+-------------------------------------------------------------------------------------------+
| First episode, currently in full remission (symptom criteria are not present)             |
+-------------------------------------------------------------------------------------------+
| Multiple episodes, currently in acute episode                                             |
+-------------------------------------------------------------------------------------------+
| Multiple episodes, currently in partial remission                                         |
+-------------------------------------------------------------------------------------------+
| Multiple episodes, currently in full remission                                            |
+-------------------------------------------------------------------------------------------+
| Continuous (subthreshold symptom periods being very brief relative to the overall course) |
+-------------------------------------------------------------------------------------------+
| Unspecified                                                                               |
+-------------------------------------------------------------------------------------------+

@@

::: {#tab:icd-11-specifier .table}
+------+--------+-----------------------------------------------------------+
| 6A25 |        | Symptomatic manifestations of primary psychotic disorders |
+------+--------+-----------------------------------------------------------+
|      | 6A25.0 | Positive symptoms in primary psychotic disorders          |
+------+--------+-----------------------------------------------------------+
|      | 6A25.1 | Negative symptoms in primary psychotic disorders          |
+------+--------+-----------------------------------------------------------+
|      | 6A25.2 | Depressive mood symptoms in primary psychotic disorders   |
+------+--------+-----------------------------------------------------------+
|      | 6A25.3 | Manic mood symptoms in primary psychotic disorders        |
+------+--------+-----------------------------------------------------------+
|      | 6A25.4 | Psychomotor symptoms in primary psychotic disorders       |
+------+--------+-----------------------------------------------------------+
|      | 6A25.5 | Cognitive symptoms in primary psychotic disorders         |
+------+--------+-----------------------------------------------------------+

@@

### 4-4-1. 현대의 아형 분류 시도

@@

::: {#tab:crow-types .table}
+---------------------------------+--------------------------------------------------------------+----------------------------------------------------------------------------+
|                                 | Type I                                                       | Type II                                                                    |
+=================================+==============================================================+============================================================================+
| Characteristic symptoms         | Delusion/hallucination, thought disorder (positive symptoms) | Affective flattening, poverty of speech, loss of drive (negative symptoms) |
+---------------------------------+--------------------------------------------------------------+----------------------------------------------------------------------------+
| Most commonly seen in           | Acute schizophrenia                                          | Chronic schizophrenia                                                      |
+---------------------------------+--------------------------------------------------------------+----------------------------------------------------------------------------+
| Drug response                   | Good                                                         | Poor                                                                       |
+---------------------------------+--------------------------------------------------------------+----------------------------------------------------------------------------+
| Outcome                         | Reversible                                                   | Irreversible                                                               |
+---------------------------------+--------------------------------------------------------------+----------------------------------------------------------------------------+
| Intellectual impairment         | Absent                                                       | Sometimes present                                                          |
+---------------------------------+--------------------------------------------------------------+----------------------------------------------------------------------------+
| Postulated pathological process | Increased dopamine receptors                                 | Cell loss and structural changes in the brain                              |
+---------------------------------+--------------------------------------------------------------+----------------------------------------------------------------------------+

>>>>

::: {#tab:dsm-5-specifier .table}
|                                                                                           |
|-------------------------------------------------------------------------------------------|
| First episode, currently in acute episode (symptom criteria are fulfilled)                |
| First episode, currently in partial remission (symptom criteria are not fully fulfilled)  |
| First episode, currently in full remission (symptom criteria are not present)             |
| Multiple episodes, currently in acute episode                                             |
| Multiple episodes, currently in partial remission                                         |
| Multiple episodes, currently in full remission                                            |
| Continuous (subthreshold symptom periods being very brief relative to the overall course) |
| Unspecified                                                                               |

@@

::: {#tab:icd-11-specifier .table}
|      |        |                                                           |
|------|--------|-----------------------------------------------------------|
| 6A25 |        | Symptomatic manifestations of primary psychotic disorders |
|      | 6A25.0 | Positive symptoms in primary psychotic disorders          |
|      | 6A25.1 | Negative symptoms in primary psychotic disorders          |
|      | 6A25.2 | Depressive mood symptoms in primary psychotic disorders   |
|      | 6A25.3 | Manic mood symptoms in primary psychotic disorders        |
|      | 6A25.4 | Psychomotor symptoms in primary psychotic disorders       |
|      | 6A25.5 | Cognitive symptoms in primary psychotic disorders         |

@@

### 4-4-1. 현대의 아형 분류 시 {#modern-subtyping}

@@

::: {#tab:crow-types .table}
|                                 | Type I                                                       | Type II                                                                    |
|---------------------------------|--------------------------------------------------------------|----------------------------------------------------------------------------|
| Characteristic symptoms         | Delusion/hallucination, thought disorder (positive symptoms) | Affective flattening, poverty of speech, loss of drive (negative symptoms) |
| Most commonly seen in           | Acute schizophrenia                                          | Chronic schizophrenia                                                      |
| Drug response                   | Good                                                         | Poor                                                                       |
| Outcome                         | Reversible                                                   | Irreversible                                                               |
| Intellectual impairment         | Absent                                                       | Sometimes present                                                          |
| Postulated pathological process | Increased dopamine receptors                                 | Cell loss and structural changes in the brain                              |

### 2022-03-18 article-01-06.Rmd
11장도 포함해서
2022-03-19

## 6-1. 적응기와 전구기

@@

::: {.l-body-outset}
+--------------------------------------------+---------------------------+--------------------------------------+-----------------------------------------+
| P. Positive                                | N. Negative               | D. Disorganization                   | G. General symptoms                     |
+============================================+===========================+======================================+=========================================+
| P1. Unusual thought content                | N1. Social anhedonia      | D1. Odd behavior or appearance       | G1. Sleep disturbance                   |
+--------------------------------------------+---------------------------+--------------------------------------+-----------------------------------------+
| P2. Suspiciousness/persecutory ideas       | N2. Avolition             | D2. Bizarre thinking                 | G2. Dysphoric mood                      |
+--------------------------------------------+---------------------------+--------------------------------------+-----------------------------------------+
| P3. Grandiose ideas                        | N3. Expression of emotion | D3. Trouble with focus and attention | G3. Motor disturbances                  |
+--------------------------------------------+---------------------------+--------------------------------------+-----------------------------------------+
| P4. Perceptual abnormalities/hallucination | N4. Experience of self    | D4. Impairment in personal hygiene   | G4. Impaired tolerance to normal stress |
+--------------------------------------------+---------------------------+--------------------------------------+-----------------------------------------+
| P5. Disorganized communication             | N5. Ideational richness   |                                      |                                         |
+--------------------------------------------+---------------------------+--------------------------------------+-----------------------------------------+
|                                            | N6. Occupational function |                                      |                                         |
+--------------------------------------------+---------------------------+--------------------------------------+-----------------------------------------+
:::

@@

## 6-3. 초발 삽화 이후

@@

::: {.l-body-outset}
+---------+-------------+---------+----------+------------+---------+---------+------+------+
| Onset   | Course      | Outcome | Lausanne | Burgholzli | Vermont | Chicago | ISoS | Mean |
+=========+=============+=========+==========+============+=========+=========+======+======+
| Acute   | Undula ting | Mild    | 25.4     | 35         | 7       | 10.8    | 29.4 | 21.5 |
+---------+-------------+---------+----------+------------+---------+---------+------+------+
| Chronic | Undula ting | Mild    | 9.6      |            | 38      | 6.8     | 22.6 | 19.3 |
+---------+-------------+---------+----------+------------+---------+---------+------+------+
| Chronic | Simple      | Severe  | 24.1     | 15         | 4       | 36.5    | 14.4 | 18.8 |
+---------+-------------+---------+----------+------------+---------+---------+------+------+
| Chronic | Undula ting | Severe  | 5.3      |            | 27      | 12.2    | 4    | 12.1 |
+---------+-------------+---------+----------+------------+---------+---------+------+------+
| Chronic | Simple      | Mild    | 10.1     | 7.5        | 12      | 4.1     | 10.4 | 8.8  |
+---------+-------------+---------+----------+------------+---------+---------+------+------+
| Acute   | Simple      | Severe  | 8.3      | 10         | 3       | 13.5    | 9.1  | 8.8  |
+---------+-------------+---------+----------+------------+---------+---------+------+------+
| Acute   | Undula ting | Severe  | 11.9     | 5          | 4       | 9.5     | 4.9  | 7.1  |
+---------+-------------+---------+----------+------------+---------+---------+------+------+
| Acute   | Simple      | Mild    | 5.3      | 5          | 5       | 6.8     | 5.3  | 5.5  |
+---------+-------------+---------+----------+------------+---------+---------+------+------+
:::

@@

## 6-4. 장기적 예후

>>>>

## 6-1. 적응기와 전구기 {#premorbid-stage}

@@

::: {.l-body-outset}
| P. Positive                                | N. Negative               | D. Disorganization                   | G. General symptoms                     |
|--------------------------------------------|---------------------------|--------------------------------------|-----------------------------------------|
| P1. Unusual thought content                | N1. Social anhedonia      | D1. Odd behavior or appearance       | G1. Sleep disturbance                   |
| P2. Suspiciousness/persecutory ideas       | N2. Avolition             | D2. Bizarre thinking                 | G2. Dysphoric mood                      |
| P3. Grandiose ideas                        | N3. Expression of emotion | D3. Trouble with focus and attention | G3. Motor disturbances                  |
| P4. Perceptual abnormalities/hallucination | N4. Experience of self    | D4. Impairment in personal hygiene   | G4. Impaired tolerance to normal stress |
| P5. Disorganized communication             | N5. Ideational richness   |                                      |                                         |
|                                            | N6. Occupational function |                                      |                                         |
:::

@@

## 6-3. 초발 삽화 이후 {#after-first-episode}

@@

::: {.l-body-outset}
| Onset   | Course      | Outcome | Lausanne | Burgholzli | Vermont | Chicago | ISoS | Mean |
|---------|-------------|---------|----------|------------|---------|---------|------|------|
| Acute   | Undula ting | Mild    | 25.4     | 35         | 7       | 10.8    | 29.4 | 21.5 |
| Chronic | Undula ting | Mild    | 9.6      |            | 38      | 6.8     | 22.6 | 19.3 |
| Chronic | Simple      | Severe  | 24.1     | 15         | 4       | 36.5    | 14.4 | 18.8 |
| Chronic | Undula ting | Severe  | 5.3      |            | 27      | 12.2    | 4    | 12.1 |
| Chronic | Simple      | Mild    | 10.1     | 7.5        | 12      | 4.1     | 10.4 | 8.8  |
| Acute   | Simple      | Severe  | 8.3      | 10         | 3       | 13.5    | 9.1  | 8.8  |
| Acute   | Undula ting | Severe  | 11.9     | 5          | 4       | 9.5     | 4.9  | 7.1  |
| Acute   | Simple      | Mild    | 5.3      | 5          | 5       | 6.8     | 5.3  | 5.5  |
:::

@@

## 6-4. 장기적 예후 {#long-term-course}
